Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-24 bbbq_article.tex
Show number of SNPs in TMHs

To statistically test whether SNPs occur just as likely in soluble as
in transmembrane protein domains, we used a Poisson binomial
distribution.
We calculated the number of SNPs that are expected to be
found in TMHs by chance, $E(n_{SNP\_TMH})$, when including
and excluding soluble proteins (with symbols
$E(n_{SNP\_TMH}^{+})$ and $E(n_{SNP\_TMH}^{-})$ respectively).
The number of SNPs that are expected to be
found in TMHs by chance when including soluble proteins,
$E(n_{SNP\_TMH}^{+})$ 
equals the number of (all) SNPs, $n_{SNP}^{+} = 38,882$, multiplied by the fraction
of all proteins that are TMH, $f_{TMH}^{+} = 9.96\%$,
resulting in 
$E(n_{SNP\_TMH}^{+}) = n_{SNP}^{+} \cdot f_{TMH}^{+} \approx 3873$
The number of SNPs that are expected to be
found in TMHs by chance when excluding soluble proteins,
$E(n_{SNP\_TMH}^{-})$ 
equals the number of SNPs in TMHs, $n_{SNP}^{-} = 21,738$, 
multiplied by the fraction of the TMPs that are TMH, $f_{TMH}^{-} = 18.66\%$,
$E(n_{SNP\_TMH}^{-}) = n_{SNP}^{-} \cdot f_{TMH}^{-} \approx 4056$.
The discrepancy between the two expectations
is caused by the distribution of SNPs over the soluble and
transmembrane proteins.
See figure \ref{fig:f_tmh_ncbi} to see which percentages
are TMH of the transmembrane proteins.
See figure HIERO to see the distribution of SNPs over the
soluble and transmembrane proteins.

>>>>

To statistically test whether SNPs occur just as likely in soluble as
in transmembrane protein domains, we used a Poisson binomial
distribution.
We calculated the number of SNPs that are expected to be
found in TMHs by chance, $E(n_{SNP\_TMH})$, when including
and excluding soluble proteins (with symbols
$E(n_{SNP\_TMH}^{+})$ and $E(n_{SNP\_TMH}^{-})$ respectively).
The number of SNPs that are expected to be
found in TMHs by chance when including soluble proteins,
$E(n_{SNP\_TMH}^{+})$ 
equals the number of (all) SNPs, $n_{SNP}^{+} = 38,882$, multiplied by the fraction
of all proteins that are TMH, $f_{TMH}^{+} = 9.96\%$,
resulting in 
$E(n_{SNP\_TMH}^{+}) = n_{SNP}^{+} \cdot f_{TMH}^{+} \approx 3873$
The number of SNPs that are expected to be
found in TMHs by chance when excluding soluble proteins,
$E(n_{SNP\_TMH}^{-})$ 
equals the number of SNPs in TMHs, $n_{SNP}^{-} = 21,738$, 
multiplied by the fraction of the TMPs that are TMH, $f_{TMH}^{-} = 18.66\%$,
$E(n_{SNP\_TMH}^{-}) = n_{SNP}^{-} \cdot f_{TMH}^{-} \approx 4056$.
The discrepancy between the two expectations
is caused by the distribution of SNPs over the soluble and
transmembrane proteins: 
there are 38,882 SNPs found in 4811 proteins (i.e. $\approx 8.08$ SNP per protein),
of which 21,738 SNPs in 2568 integral membrane proteins (i.e. $\approx 8.46$
SNP per integral membrane protein).

\richel{How many SNPs in TMHs?}
See figure \ref{fig:f_tmh_ncbi} to see which percentages
are TMH of the transmembrane proteins.
